### 2013-10-29 tex/2-related-work.tex
Snapshot of the project as sent in the first email.

\section{Related work}
2-related-work.tex

Web Object Identification for Web Automation and Meta-Search,
\surl{http://www.dbai.tuwien.ac.at/proj/tamcrow/download/Kordomatis2013WIMS.pdf}

Self-supervised Automated Wrapper Generation for Weblog Data Extraction,
\surl{https://github.com/OlivierBlanvillain/blogforever-crawler-publication/raw/master/papers/bncod_published.pdf}

Web Data Extraction, Applications and Techniques: A Survey,
\surl{http://www.emilio.ferrara.name/wp-content/uploads/2011/07/survey-csur.pdf}

Archiving Data Objects using Web Feeds,
\surl{http://hal.archives-ouvertes.fr/docs/00/53/79/62/PDF/iwawienna.pdf}

Intelligent and Adaptive Crawling of Web Applications for Web Archiving,
\surl{http://pierre.senellart.com/publications/faheem2013intelligent.pdf}

Zero-cost Labelling with Web Feeds for Weblog Data Extraction,
\surl{http://www2013.org/companion/p73.pdf}

OXPath: A Language for Scalable, Memory-efficient Data Extraction from Web Applications,
\surl{http://www.vldb.org/pvldb/vol4/p1016-furche.pdf}