### 2013-10-29 tex/3-architecture.tex

\section{Architecture}

\subsection{Scrapy and Invenio}
Introduce both projects with their respective strengths, tool/technology justification.

\subsection{JavaScript Rendering}
\begin{enumerate}
  \item PhantomJS
  \item "Show more comments" click
  \item Done upfront, every thing else is independent
\end{enumerate}

\subsection{Extraction Process}
High level overview: XPath build from RSS feeds