### 2013-10-29 tex/4-algorithms.tex

\section{Algorithms}

\subsection{Content extraction}
\begin{enumerate}
  \item The algorithm
  \item Choice of selectors (id, else class)
  \item Choice of string similarity
\end{enumerate}

\subsection{Variations for comments and date}
All the details about how to use the same algorithm to extract comments and dates.