### 2013-11-01 tex/3-architecture.tex
Improve score from 68.2% to 73.7% by taking top queries for all pages for the feed and not only the first one.

\section{Architecture}

\subsection{Scrapy and Invenio}
Introduce both projects with their respective strengths, tool/technology justification.

\begin{enumerate}
  \item Stateless
  \item Trade of efficiency for addaptability
  \item Stuctural maintenence is given because by recompute them each time
\end{enumerate}

\subsection{JavaScript Rendering}
\begin{enumerate}
  \item PhantomJS
  \item "Show more comments" click
  \item Done upfront, every thing else is independent
\end{enumerate}

While some (Archiving Data Objects using Web Feeds -> Perenniality of Web archives.) take the

Google Blogspot Dynamic Views: \surl{http://buzz.blogger.com/2012/02/dynamic-views-update-3-gadgets.html}, \surl{https://support.google.com/blogger/answer/1229061?hl=en}

\subsection{Extraction Process}
High level overview: XPath build from RSS feeds