### 2013-11-04 tex/0-abstract.tex

Blogs are by nature a very dynamic and volatile communication medium. The
BlogForever project aims to provide a generic platform for blog preservation,
thus ensuring access to this information to future generations. This paper
presents a key component of this archiving system: the web crawler. More
precisely, our discussion will concentrate on techniques to automatically
extract content, authors, date and comments of blog posts. One of the main
challenges in the preservation of web objects is the rapid evolution of
technologies. For instance, the recent trend towards using JavaScript to
dynamically generate pages on the client side constitutes a new obstacle for
today's crawlers. We will show how we integrated a web browser into the
harvesting process in order to address these matters. Furthermore, we present a
simple and robust algorithm based on string matching between web feeds and page
content to generate wrappers. Being applicable to all posts of a blog, the
resulting extraction procedure scales for large blogs.