### 2013-11-06 tex/4-architecture.tex
Algorithms assumptions.

\section{Architectute}

This section gives an overview of the overall software architecture of our crawler. After a breif introductions of the tool, we will explain how our approach allow to use the crawler on a large scale distributed architectute. We will then see how we integrated a headless web browser into the harvesting process to support blogs that use JavaScript to dynamically generate post content and comments.

\subsection{Architecture}
- scrapy, dl policies and post identification \\
- item pipeline, modularity \\
- js rendering done upfront, everything else general \\

\subsection{Scalability}
- stateless \\
- scrapyd \\

\subsection{JavaScript rendering}
- challenges \\
Ajax
Google Blogspot Dynamic Views: \surl{http://buzz.blogger.com/2012/02/dynamic-views-update-3-gadgets.html}, \surl{https://support.google.com/blogger/answer/1229061?hl=en}
- present phjs, also here screenshot \\
- ez of use vs power \\
- our click show more use case \\
- parallelization \\