### 2013-11-25 tex/2-related-work.tex
Minor fixes in captions + some of variations subsection.

\section{Related work}
2-related-work.tex

Web Object Identification for Web Automation and Meta-Search,
\\
http://www.dbai.tuwien.ac.at/proj/ta\\
mcrow/download/Kordomatis2013WIMS.pdf

Self-supervised Automated Wrapper Generation for Weblog Data Extraction,
\\
https://github.com/OlivierBlanvillai\\
n/blogforever-crawler-publication/raw\\
/master/papers/bncod\_published.pdf

Web Data Extraction, Applications and Techniques: A Survey,
\\
http://www.emilio.ferrara.name/wp-co\\
ntent/uploads/2011/07/survey-csur.pdf

Archiving Data Objects using Web Feeds,
\\
http://hal.archives-ouvertes.fr/docs\\
/00/53/79/62/PDF/iwawienna.pdf

Intelligent and Adaptive Crawling of Web Applications for Web Archiving,
\\
http://pierre.senellart.com/publicat\\
ions/faheem2013intelligent.pdf

Zero-cost Labelling with Web Feeds for Weblog Data Extraction,
\\
http://www2013.org/companion/p73.pdf\\

OXPath: A Language for Scalable, Memory-efficient Data Extraction from Web Applications,
\\
http://www.vldb.org/pvldb/vol4/p1016\\
-furche.pdf