### 2013-11-27 tex/testfigure.tex
Add first gnuplot figure.

\begin{figure}[ht]
  \centering
  \begin{gnuplot}[scale=.75,terminal=epslatex,terminaloptions=color]
    set size ratio 0.618
    set yrange [0:50]
    set xrange [0:70]
    plot "data/runningtime.txt"\
       using 0:1 smooth cumulative with lines linewidth 5 title 'ContentExtractor',\
    "" using 0:2 smooth cumulative with lines linewidth 5 title 'Boilerpipe',\
    "" using 0:3 smooth cumulative with lines linewidth 5 title 'Goose',\
    "" using 0:4 smooth cumulative with lines linewidth 5 title 'Readability'
  \end{gnuplot}
\end{figure}
%lskdfjlsjflskjfls