### 2013-11-28 tex/testfigure.tex
Updates after discussion + ongoing evaluation.

\begin{figure}[ht]
  \centering
  \begin{gnuplot}
    set terminal epslatex color
    set size 0.71
    set size ratio 0.618

    # set title 'title'
    set ylabel 'Cumulated running time (sec.)' offset 1.3
    set xlabel 'Processed pages'

    set yrange [0:55]
    set xrange [0:70]

    set arrow from 15,graph(0,0) to 15,graph(1,1) nohead linecolor rgb 'grey'

    plot 'data/runningtime.txt'\
       u 0:1 smooth cumulative w lines linewidth 5 title 'Our Approach',\
    '' u 0:4 smooth cumulative w lines linewidth 5 title 'Readability',\
    '' u 0:2 smooth cumulative w lines linewidth 5 title 'Boilerpipe',\
    '' u 0:3 smooth cumulative w lines linewidth 5 title 'Goose',\
  \end{gnuplot}
\end{figure}
%lskdfjlsjflskjfls