### 2013-12-05 tex/0-abstract.tex
Minor fixes, add main.pdf under version control.

Blogs are a dynamic communication medium which has been widely established on the web. The BlogForever project has developed an innovative system to harvest, preserve, manage and reuse blog content. This paper presents a key component of the BlogForever platform, the web crawler. More precisely, our work concentrates on techniques to automatically extract content such as articles, authors, dates and comments from blog posts. To achieve this goal, we introduce a simple and robust algorithm to generate extraction rules based on string matching using the blog's web feed in conjunction to blog hypertext. This approach is leading to a scalable blog data extraction process. Furthermore, we show how we integrate a web browser into the web harvesting process in order to support the data extraction from blogs with JavaScript generated content.