### 2014-01-14 tex/2-related-work.tex
Improvements in the Related work section

\section{Related work}

Web archiving is a challenging and very interesting problem and has been the subject of many studies as well as the goal of several projects.
\comment{Mention Internet Archive here and maybe another generic web archiving project}

A fraction of web archiving studies and projects focus on the specific problem of blog harvesting and archiving.
They focus on the unique nature of blogs as dynamic web resources with a usually common set of attributes.
Each entry in a blog, commonly called a post, has a title, an author, a publication date, some content and possibly some descriptive tags.
(In this section we review some studies and projects that introduce and describe such techniques.)

Blogs are strongly associated with web feeds, using technologies such as RSS and Atom (expand and reference - needed?).
Web feeds are structured documents (often XML-based) that advertize and provide access to a website's content.
Given their format, web feeds constitute an excellent source for blog archiving.
They do, however, present some problems: (a) web feeds contain only the latest posts of a blog (usually 20) and (b) web feeds may only provide a summary or a fraction of the entire content of each blog post.
More advanced methods and techniques have to be employed in order to achieve quality blog archiving.

Here mention projects that use web feeds for discovery and compare them to ours.
\begin{itemize}
  \item ( Zero-cost Labelling with Web Feeds for Weblog Data Extraction - http://www2013.org/companion/p73.pdf).
  \item ( Self-supervised Automated Wrapper Generation for Weblog Data Extraction - https://github.com/OlivierBlanvillain/blogforever-crawler-publication/raw/master/papers/bncod\_published.pdf).
  \item ( Archiving Data Objects using Web Feeds - http://hal.archives-ouvertes.fr/docs/00/53/79/62/PDF/iwawienna.pdf).
\end{itemize}
Some of these should also be in the comparison for extraction methods.

Here mention projects that use other techniques for discovery and compare them to ours.
\begin{itemize}
  \item ???
\end{itemize}

Here mention projects that use various techniques for extraction and compare them to ours.
\begin{itemize}
  \item ( Web Object Identification for Web Automation and Meta-Search - http://www.dbai.tuwien.ac.at/proj/tamcrow/download/Kordomatis2013WIMS.pdf).
  \item ( OXPath: A Language for Scalable, Memory-efficient Data Extraction from Web Applications - http://www.vldb.org/pvldb/vol4/p1016-furche.pdf).
  \item ( Intelligent and Adaptive Crawling of Web Applications for Web Archiving - http://pierre.senellart.com/publications/faheem2013intelligent.pdf).
  \item ( Web Data Extraction, Applications and Techniques: A Survey - http://www.emilio.ferrara.name/wp-content/uploads/2011/07/survey-csur.pdf).
\end{itemize}

Here mention that there is no comparison for the JavaScript rendering as there are no similar projects.
\begin{itemize}
  \item ???
\end{itemize}