### 2014-01-19 tex/main.tex

% \documentclass[draft]{acm_proc_article-sp}
\documentclass{acm_proc_article-sp}

\usepackage[ruled, vlined]{algorithm2e}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[autostyle=false, style=english]{csquotes}
\usepackage[usenames,dvipsnames]{color}
\usepackage[justification=centering]{caption}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{gnuplottex}
\usepackage{pifont}
\usepackage{quoting}
\usepackage{lmodern}
\usepackage[hyphens]{url}
\usepackage{sepfootnotes}
\usepackage{setspace}
\usepackage[pdfpagelabels=false]{hyperref}
\usepackage{hypcap}

\MakeOuterQuote{"}
\newfootnotes{A}
\quotingsetup{vskip=0pt,font={itshape}}
\setenumerate{noitemsep,topsep=0pt,parsep=2pt,partopsep=0pt}
\setitemize{noitemsep,topsep=0pt,parsep=4pt,partopsep=0pt}
\setlength\floatsep{1.25\baselineskip plus 3pt minus 2pt}
\setlength\textfloatsep{1.25\baselineskip plus 3pt minus 2pt}
\setlength\intextsep{1.25\baselineskip plus 3pt minus 2 pt}
\def\algorithmautorefname{Algorithm}
\graphicspath{{img/}}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand\TODO[1]{\emph{\color{Red} TODO: #1}}
\renewcommand\comment[1]{\emph{\color{Gray}#1}}
\newcommand\code[1]{\texttt{#1}}
\newcommand\coderef[1]{\texttt{\ref{#1}}}
\newcommand\stringliteral[1]{\texttt{\char`\"#1\char`\"}}
\newcommand\len[1]{|#1|}
\newcommand\set[1]{\{#1\}}
\newcommand\surl[1]{\urlstyle{ff}\url{#1}}
\newcommand\Oof[1]{$\mathcal{O}(#1)$}

\begin{document}

  \title{BlogForever Crawler: Techniques and Algorithms\\to Harvest Modern Weblogs}

  \numberofauthors{3}

  \author{
  % 1st. author
  \alignauthor
  Olivier Blanvillain\\
         \affaddr{École polytechnique fédérale de Lausanne (EPFL)}\\
         \affaddr{1015 Lausanne, Switzerland}\\
         \email{\surl{olivier.blanvillain@epfl.ch}}
  % 2nd. author
  \alignauthor
  Nikos Kasioumis\\
         \affaddr{European Organization for Nuclear Research (CERN)}\\
         \affaddr{1211 Geneva 23, Switzerland}\\
         \email{\surl{nikos.kasioumis@cern.ch}}
  % 3rd. author
  \alignauthor
  Vangelis Banos\\
         \affaddr{Department of Informatics}\\
         \affaddr{Aristotle University of Thessaloniki, Greece}\\
         \email{\surl{vbanos@gmail.com}}
  }

  \hypersetup{pageanchor=false}
  \maketitle

  \begin{abstract}
    \input{0-abstract.tex}
  \end{abstract}

  \category{H.3.3}{Information Storage and Retrieval}{Information Search and Retrieval}[Information filtering, Query formulation, Selection process]
  \category{\\D.2.8}{Software Engineering}{Metrics}[Complexity measures, Performance measures]

  % In order of most significant, first letter of each word capitalized.
  \terms{Design, Algorithms, Performance, Experimentation}

  % In alphabetical order, first letter of the first word capitalized.
  \keywords{Blog crawler, web data extraction, wrapper generation}

  \input{algos.tex}
  \input{tables.tex}
  ~\\~\\~\\~\\~\\~\\~\\
  \input{1-intro.tex}
  \clearpage
  \input{2-algorithms.tex}
  \clearpage
  \input{3-architecture.tex}
  \clearpage
  \input{4-evaluation.tex}
  \clearpage
  \input{5-related-work.tex}
  \clearpage
  \input{6-conclusion-and-future-work.tex}

  \bibliographystyle{abbrv}
  \bibliography{main,manual}

\end{document}