### 2014-04-01 tex/runningtimeFigure.tex

\begin{figure}[t]
  \capstart
  \refstepcounter{figure}
  \label{runningtime}
  \vspace{-12pt}
  \hspace{-34pt}
  \begin{gnuplot}% nop
    set terminal epslatex color
    set size 0.76
    set size ratio 0.618

    set title 'Figure \arabic{figure}: Running time of articles extraction.'
    set ylabel 'Cumulated running time (sec.)' offset 1.3
    set xlabel 'Processed blog posts'

    set yrange [0:44]
    set xrange [0:65]

    set arrow from 15,graph(0,0) to 15,graph(1,1) nohead linecolor rgb 'grey' linetype 2

    plot 'data/runningtime.txt'\
       u 0:1 smooth cumulative w lines linewidth 3 title 'Our approach',\
    '' u 0:4 smooth cumulative w lines linewidth 4 title 'Readability',\
    '' u 0:2 smooth cumulative w lines linewidth 4 title 'Boilerpipe',\
    '' u 0:3 smooth cumulative w lines linewidth 4 title 'Goose'
  \end{gnuplot}
  \vspace{-12pt}
\end{figure}