### 2014-04-25 tex/main.tex
Update license

% \documentclass[draft]{sig-alternate}
\documentclass{sig-alternate}

\parindent 0pt
\usepackage[ruled, vlined]{algorithm2e}
\parindent 10pt
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[autostyle=false, style=english]{csquotes}
\usepackage[usenames,dvipsnames]{color}
\usepackage[nodayofweek]{datetime}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage[tableposition=top, font=bf]{caption}
\usepackage{balance}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{gnuplottex}
\usepackage{ifthen}
\usepackage{quoting}
\usepackage{sepfootnotes}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage[hyphens]{url}
\usepackage[pdfpagelabels=false,draft=true]{hyperref}
\usepackage{hypcap}

\MakeOuterQuote{"}
\def\algorithmautorefname{Algorithm}

\newdateformat{lv}{\THEDAY{ }\monthname[\THEMONTH] \THEYEAR}
\newcommand{\lastvisited}{Last visited \lv\today}
\newcommand\TODO[1]{\emph{\color{Red} TODO: #1}}
\newcommand\noop[1]{\emph{\emph{#1}}}
\newcommand\edit[1]{\emph{\color{Green}#1}}
\newcommand\code[1]{\texttt{#1}}
\newcommand\coderef[1]{\texttt{\ref{#1}}}
\newcommand\stringliteral[1]{\texttt{\char`\"#1\char`\"}}
\newcommand\len[1]{|#1|}
\newcommand\set[1]{\{#1\}}
\newcommand\surl[1]{\urlstyle{ff}\url{#1}}
\newcommand\Oof[1]{$\mathcal{O}(#1)$}
\newboolean{fref}

\begin{document}

  \title{BlogForever Crawler: Techniques and Algorithms\\to Harvest Modern Weblogs}

  \numberofauthors{3}

  \author{
  \alignauthor
  Olivier Blanvillain\\
         \affaddr{École Polytechnique Fédérale de Lausanne (EPFL)}\\
         \affaddr{1015 Lausanne, Switzerland}\\
         \email{\href{mailto:olivier.blanvillain@epfl.ch}{olivier.blanvillain@epfl.ch}}
  \alignauthor
  Nikos Kasioumis\\
         \affaddr{European Organization for Nuclear Research (CERN)}\\
         \affaddr{1211 Geneva 23, Switzerland}\\
         \email{\href{mailto:nikos.kasioumis@cern.ch}{nikos.kasioumis@cern.ch}}
  \alignauthor
  Vangelis Banos\\
         \affaddr{Department of Informatics}\\
         \affaddr{Aristotle University of Thessaloniki, Greece}\\
         \email{\href{mailto:vbanos@gmail.com}{vbanos@gmail.com}}
  }

  \permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than the author(s) must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from Permissions@acm.org.}

  \conferenceinfo{WIMS'14,}{June 02-04, 2014 Thessaloniki, Greece}
  \crdata{978-1-4503-2538-7/14/06}
  \newcommand{\doi}{http://dx.doi.org/10.1145/2611040.2611067}

  \global\copyrightetc{Copyright is held by the owner/authors. Publication rights licensed to ACM. ACM \the\acmcopyr\$15.00.\\\doi}

  \maketitle

  \begin{abstract}
    \input{0-abstract.mdtex}
  \end{abstract}

  \category{H.3.3}{Information Storage and Retrieval}{Information Search and Retrieval}[Information filtering, Query formulation, Selection process]
  \category{\\D.2.8}{Software Engineering}{Metrics}[Complexity measures, Performance measures]
  \terms{Design, Algorithms, Performance, Experimentation}
  \keywords{Blog crawler, web data extraction, wrapper generation}

  \input{algos.tex}
  \input{tables.tex}

  \section{Introduction}
  \label{introduction}
  \input{1-introduction.mdtex}

  \section{Algorithms}
  \label{algorithms}
  \input{2-algorithms.mdtex}

  \section{Architecture}
  \label{architecture}
  \input{3-architecture.mdtex}

  \section{Evaluation}
  \label{evaluation}
  \input{4-evaluation.mdtex}

  \section{Related Work}
  \label{relatedwork}
  \input{5-related-work.mdtex}

  \section{Conclusion and Future Work}
  \label{conclusionandfuturework}
  \input{6-conclusion-and-future-work.mdtex}

  \section{Acknowledgments}
  \label{acknowledgments}
  \input{7-acknowledgments.mdtex}

  \vspace{3pt}

  \balance

  \def\changemargin{\list{}{\leftmargin3pt}\item[]}
  \let\endchangemargin=\endlist
  \let\oldbibitem\bibitem

  \setboolean{fref}{true}
  \renewcommand\bibitem[1]{
    \ifthenelse{\boolean{fref}}{
      \setboolean{fref}{false}
      \vspace{4pt}
      \oldbibitem{#1}
      \begin{changemargin}
    }{
      \end{changemargin}
      \oldbibitem{#1}
      \begin{changemargin}
    }
  }

  \bibliographystyle{abbrv}
  \input{hardcoded.bbl}
  % \bibliography{short}

\end{document}