### 2014-12-16 0-abstract.tex
As of Sep 30, 2014 diff --git a/0-abstract.tex b/0-abstract.tex new file mode 100644 index 0000000..642c9e6 --- /dev/null +++ b/0-abstract.tex @@ -0,0 +1,18 @@ + +Blogs are one of the most prominent means of communication on the web. +Their content, interconnections and influence constitute a unique +socio-technical artefact of our times which needs to be preserved. +The BlogForever project has established best practices and developed an +innovative system to harvest, preserve, manage and reuse blog content. +This paper presents the latest developments of the blog crawler which +is a key compo

Blogs are one of the most prominent means of communication on the web. 
Their content, interconnections and influence constitute a unique 
socio-technical artefact of our times which needs to be preserved. 
The BlogForever project has established best practices and developed an 
innovative system to harvest, preserve, manage and reuse blog content. 
This paper presents the latest developments of the blog crawler which 
is a key component of the BlogForever platform. More precisely, our 
work concentrates on techniques to automatically extract content such 
as articles, authors, dates and comments from blog posts. To achieve 
this goal, we introduce a simple yet robust and scalable algorithm to 
generate extraction rules based on string matching using the blog's 
web feed in conjunction with blog hypertext. Furthermore, we present a 
system architecture which is characterised by efficiency, modularity, 
scalability and interoperability with third-party systems. Finally, 
we conduct thorough evaluations of the performance and accuracy of
our system.