### 2014-12-16 3-algorithms.tex

\section{Algorithms}
\label{algorithms}
We propose a new set of algorithms to extract blog post articles as well 
as variations for extracting authors, dates and comments. We start with 
our motivation to use blog specific characteristics, followed by our 
approach to build \emph{extraction rules} which are applicable throughout 
a blog. Our focus is on minimising the algorithmic complexity while 
keeping our approach simple and generic.

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation}
\label{motivation}

Extracting metadata and content from HTML documents is a challenging task 
because standards and format recommendations suffer from very low usage. 
W3C has been publishing web standards and format recommendations for 
quite some time~\cite{w3c2014}. For instance, according to the W3C 
HTML guidelines, the \code{<h1></h1>} tags have to contain the 
highest-level heading of the page and must not appear more than 
once per page~\cite{w3c2002}. More recently, specifications such 
as microdata~\cite{whatwg2014} define ways to embed semantic information 
and metadata inside HTML documents, but these still suffer from 
very low usage: estimated to be used in less than 0.5\% of 
websites~\cite{andrewrogers2014}. In fact, the majority of websites rely 
on the generic \code{<span></span>} and \code{<div></div>} container 
elements with custom \code{id} or \code{class} attributes to organise 
the structure of pages, and more than 95\% of pages do not pass HTML 
validation~\cite{brianwilson2008}. Under such circumstances, relying 
on HTML structure to extract content from web pages is not viable and 
other techniques need to be employed.

Having blogs as our target websites, we made the following observations 
which play a central role in the extraction process:

\begin{enumerate}
\item\label{havefeedAssum} 
Blogs provide web feeds: structured and standardized views of the latest 
posts of a blog.
\item\label{similarhtmlAssum} 
Posts of the same blog share a similar HTML structure.
\end{enumerate}

Web feeds usually contain about 20 blog posts~\cite{oita2010}, 
often less than the total number of posts in blogs. Consequently, to 
effectively archive the entire content of a blog, it is necessary to 
download and process pages beyond the ones referenced in the web feed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Content extraction overview}
\label{contentextraction}

% per blog procedure, related work
To extract content from blog posts, we proceed by building 
\emph{extraction rules} from the data given in the blog's web feed. 
The idea is to use a set of \emph{training data}, pairs of HTML pages 
and target content, which are used to build an extraction rule capable 
of locating the target content on each HTML page.

% input from observations.
Observation (\ref{havefeedAssum}) allows the crawler to obtain input for 
the extraction rule generation algorithm: each web feed entry contains a 
link to the corresponding web page as well as blog post article (either 
abstract or full text), its title, authors and publication date. We call 
these fields \emph{targets} as they constitute the data our crawler aims 
to extract. Observation (\ref{similarhtmlAssum}) guarantees the existence 
of an appropriate extraction rule, as well as its applicability to all 
posts of the blog.

\autoref{extractionAlgo} shows the generic procedure we use to build 
extraction rules. The idea is quite simple: for each \code{(}\emph{page, 
target}\code{)} input, compute, out of all possible extraction rules, 
the best one with respect to a certain \code{ScoreFunction}. The rule 
which is most frequently the \emph{best rule} is then returned.

\extractionAlgo

One might notice that each \emph{best rule} computation is 
independent and operates on a different input pair. This implies that 
\autoref{extractionAlgo} is \emph{embarrassingly parallel}: iterations 
of the outer loop can trivially be executed on multiple threads.

Functions in \autoref{extractionAlgo} are voluntarily abstract at this 
point and will be explained in detail in the remaining of this section.
\hyperref[extractionrulesandstringsimilarity]{Subsection~\ref*{extractionrulesandstringsimilarity}} % Waiting for \Autoref{}...
defines \code{AllRules}, \code{Apply} and the \code{ScoreFunction} we use 
for article extraction. In \autoref{timecomplexityandlinearreformulation} 
we analyse the time complexity of \autoref{extractionAlgo} and give a 
linear time reformulation using dynamic programming. Finally, 
\autoref{variationsforauthorsdatesandcomments} shows how the 
\code{ScoreFunction} can be adapted to extract authors, dates and 
comments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extraction rules and string similarity}
\label{extractionrulesandstringsimilarity}

In our implementation, rules are queries in the XML Path Language 
(XPath). Consequently, standard libraries can be used to parse HTML pages 
and apply extraction rules, providing the \code{Apply} function used in 
\autoref{extractionAlgo}. We experimented with 3 types of XPath queries: 
selection over the HTML \code{id} attribute, selection over the HTML 
\code{class} attribute and selection using the relative path in the HTML 
tree. \code{id} attributes are expected to be unique, and \code{class} 
attributes have showed in our experiments to have better consistency than 
relative paths over pages of a blog. For these reasons we opt to always 
favour \code{class} over path, and \code{id} over \code{class}, such 
that the \coderef{allrulesAlgo} function returns a single rule per node.

\allrulesAlgo

Unsurprisingly, the choice of \code{ScoreFunction} greatly influences 
the running time and precision of the extraction process. When targeting 
articles, extraction rule scores are computed with a string similarity 
function comparing the extracted strings with the target strings. 
We chose the Sorensen–Dice coefficient similarity~\cite{dice1945}, 
which is, to the best of our knowledge, the only string similarity 
algorithm fulfilling the following criteria:

\begin{itemize}
\item\label{wordorderProp} Has low sensitivity to word ordering,
\item\label{lengthProp} Has low sensitivity to length variations,
\item\label{linearProp} Runs in linear time.
\end{itemize}

Properties \ref{wordorderProp} and \ref{lengthProp} are essential when 
dealing with cases where the blog's web feed only contains an abstract 
or a subset of the entire post article. \autoref{similarityTable} 
gives examples to illustrate how these two properties hold for the 
Sorensen–Dice coefficient similarity but do not for \emph{edit distance} 
based similarities such as the Levenshtein~\cite{levenshtein1966} 
similarity.

\begin{table}
\tbl{Examples of string similarities.}
{\begin{tabular}{@{}llcc@{}} \toprule
        \textbf{string1} & \textbf{string2} & \textbf{Sorensen-Dice} & \textbf{Levenshtein} \\
    \midrule
    \stringliteral{Scheme Scala} & \stringliteral{Scala Scheme}
    & 90\% & 50\% \\
    \stringliteral{Rachid} & \stringliteral{Richard}
    & 18\% & 61\% \\
    \stringliteral{Rachid} & \stringliteral{Amy, Rachid and all their friends}
    & 29\% & 31\% \\ \hline
\end{tabular}}
\label{similarityTable}
\end{table}

The Sorensen–Dice coefficient similarity algorithm operates by 
first building sets of pairs of adjacent characters, also known as 
\emph{bigrams}, and then applying the \emph{quotient of similarity} formula:

\similarityAlgo

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time complexity and linear reformulation}
\label{timecomplexityandlinearreformulation}

With the functions \coderef{allrulesAlgo}, \code{Apply} and 
\coderef{similarityAlgo} (as \code{ScoreFunction}) being defined, the 
definition of \autoref{extractionAlgo} for article extraction is now 
complete. We can therefore proceed with a time complexity analysis.

First, let's assume that we have at our disposal a linear time HTML 
parser that constructs an appropriate data structure, indexing HTML 
nodes on their \code{id} and \code{class} attributes, effectively making 
\code{Apply}~$\in$~\Oof{1}. As stated before, the outer loop splits 
the input into independent computations and each call to 
\coderef{allrulesAlgo} returns (in linear time) at most as many rules as 
the number of nodes in its \emph{page} argument. Therefore, the body of 
the inner loop will be executed \Oof{n} times. Because each extraction 
rule can return any subtree of the queried page, each call to 
\coderef{similarityAlgo} takes \Oof{n}, leading to an overall quadratic 
running time.

We now present \autoref{linearAlgo}, a linear time reformulation of 
\autoref{extractionAlgo} for article extraction using dynamic programming.

\linearAlgo

While very intuitive, the original idea of first generating extraction 
rules and then picking these best rules prevents us from effectively 
reusing previously computed $N$-grams (sets of adjacent characters). 
For instance, when evaluating the extraction rule for the HTML root node, 
\autoref{extractionAlgo} will obtain the complete string of the page 
and pass it to the \coderef{similarityAlgo} function. At this point, 
the information on where the string could be split into substrings with 
already computed $N$-grams is not accessible, and the $N$-grams of 
the page have to be computed by linearly traversing the entire string. 
To overcome this limitation and implement \emph{memoization} over the 
$N$-gram computations, \autoref{linearAlgo} uses a post-order traversal 
of the HTML tree and computes node $N$-grams from their children 
$N$-grams. This way, we avoid serializing HTML subtrees for each 
$N$-gram computation and have the guarantee that each character of the 
HTML page will be read at most once during the $N$-gram computation.

An interesting question is what is the optimal $N$ for $N$-gram 
computation. To answer this question, we conduct some simple experiments. 
Using sample blog post excerpts and full texts, we calculate the string 
similarity between each excerpt and full text. For instance, using a blog 
post full text and associated text excerpt from the popular blog TechCrunch
as presented in Table \ref{table:techcrunch}, the text similarity results
(Table \ref{table:similarity-n-values}) using different $N$ values support
the selection of bigrams ($N$=2).

\begin{table}
\tbl{TechCrunch blog post example.}
{\begin{tabular}{@{}p{12cm}@{}} \toprule
\emph{Blog post text} \\ \hline
Apple has a new version of iOS 8 out, just a short time after the 
initial launch of the software. The update, 8.0.1, includes a number of 
fixes, but most notably (and listed first), it addressed the bug that 
prevented HealthKit apps from being available at launch. It also zaps 
some bugs with third-party keyboards, which should make them remain the 
default option until a user switches to another, which has been a sore 
spot for fans of the new external software keyboard options.

Unfortunately, installing the iOS 8.0.1 update revealed that despite 
Apple’s promised fixes, it actually completely disables cellular service 
and Touch ID on many devices, though some iPhone 5s and older model 
owners report no issues. The bottom line is that you should definitely 
NOT install this update, at least until an updated version appears, 
at which time we’ll let you know it’s safe to go ahead.

As you can see in the image below, Apple is also addressing an issue that 
blocked some photos from appearing in Photo Library, fixing reliability 
concerns around Reachability on iPhone 6 and 6 Plus (which brings the top 
of the screen down when you double touch the Home button), zapping bugs 
that cause unexpected data use when communicating via SMS or MMS, and 
improving the “Ask to Buy” feature for Family Sharing, specifically 
around in-app purchases, in addition to other minor bugs.\\ \hline

\emph{Weblog post text abstract from RSS} \\ \hline
Apple has a new version of iOS 8 out, just a short time after the 
initial launch of the software. The update, 8.0.1, includes a number of 
fixes, but most notably (and listed first), it addressed the bug that 
prevented HealthKit apps from being available at launch. It also zaps... \\ \hline
\end{tabular}}
\label{table:techcrunch}
\end{table}

\begin{table}
\tbl{Blog post excerpt and full text similarity using different $N$ values.}
{\begin{tabular}{@{}lccc@{}} \toprule
        \emph{$N$} & \emph{2} & \emph{3} & \emph{4} \\ \hline
Similarity & 0.7817 & 0.6680 & 0.6212 \\ \hline
\end{tabular}}
\label{table:similarity-n-values}
\end{table}

With bigrams computed in this dynamic programming manner, the overall 
time to compute all \coderef{bigramsAlgo}\code{(}\emph{node.text}\code{)} 
is linear. To conclude the proof that \autoref{linearAlgo} runs in linear 
time we show that all other computations of the inner loop can be done 
in constant \emph{amortized} time. As the number of edges in a tree is 
one less than the number of nodes, the \emph{amortized} number of bigrams 
unions per inner loop iteration tends to one. Each \emph{quotient of 
similarity} computation requires one bigrams intersection and three 
bigrams length computations. Over a finite alphabet (we used printable 
ASCII), bigrams sizes have bounded size and each of these operations 
takes constant time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Variations for authors, dates, comments}
\label{variationsforauthorsdatesandcomments}

Using string similarity as the only score measurement leads to poor 
performance on author and date extraction, and is not suitable for 
comment extraction. This subsection presents variations of the 
\code{ScoreFunction} which addresses issues of these other types of content.

The case of authors is problematic because authors' names often appear in 
multiple places of a page, which results in several rules with maximum 
\coderef{similarityAlgo} score. The heuristic we use to get around this 
issue consists of adding a new component in the \code{ScoreFunction} for 
author extraction rules: the \emph{tree distance} between the evaluated 
node and the post content node. This new component takes advantage of 
the positioning of a post's authors node which often is a direct child 
or shares its parent with the post content node.

Dates are affected by the same duplication issue, as well as the issue of 
inconsistencies of format between web feeds and web pages. Our solution 
for date extraction extends the \code{ScoreFunction} for authors by 
comparing the \emph{extracted} string to multiple \emph{targets}, each 
being a different string representation of the original date obtained 
from the web feed. For instance, if the feed indicates that a post was 
published at \stringliteral{Thu, 01 Jan 1970 00:00:00}, our algorithm 
will search for a rule that returns one of \stringliteral{Thursday 
January 1, 1970}, \stringliteral{1970-01-01}, \stringliteral{43 years 
ago} and so on. So far we do not support dates in multiple languages, 
but adding new target formats based on languages detection would be a 
simple extension of our date extraction algorithm.

Comments are usually available in separate web feeds, one per blog post. 
Similarly to blog feeds, comment feeds have a limited number of entries, 
and when the number of comments on a blog post exceeds this limit, 
comments have to be extracted from web pages. To do so, we use the 
following \code{ScoreFunction}:

\begin{itemize}
\item 
Rules returning fewer HTML nodes than the number of comments on the feed 
are filtered out with a zero~score,
\item 
The scores of the remaining rules are computed with the value of the 
\emph{maximum weighted matching} in the \emph{complete bipartite graph} 
$G = (U, V, E)$, where $U$ is the set of HTML nodes returned by the rule, 
$V$ is the set of target comment fields from the web feed (such as 
comment authors) and $E(u, v)$ has weight equal to 
\coderef{similarityAlgo}\code{(}\emph{u, v}\code{)}.
\end{itemize}

Regarding time complexity, computing the \emph{tree distance} of each 
node of a graph to a single reference node can be done in linear time, 
and multiplying the number of targets by a constant factor does not 
affect the asymptotic computational complexity. Computing scores of 
comment extraction rules requires a more expensive algorithm. However, 
this is compensated by the fact that the proportion of candidates left, 
after filtering out rules not returning enough results, is very low in 
practice. Analogous reformulations to the one done with 
\autoref{linearAlgo} can be straightforwardly applied on each 
\code{ScoreFunction} to minimize the time spent in 
\coderef{similarityAlgo}.