### 2014-12-16 5-evaluation.tex

\section{Evaluation}
\label{evaluation}

Our evaluation is articulated in two parts. First, we compare the 
article extraction procedure presented in \autoref{algorithms} with 
three open-source projects capable of extracting articles and titles 
from web pages. The comparison will show that our weblog-targeted solution 
has better performance both in terms of success rate and running time. 
Second, a discussion is held regarding the different solutions available 
to archive data beyond what is available in the HTML source code. 
Extraction of authors, dates and comments is not part of this evaluation 
because of the lack of publicly available competing projects and 
reference data sets.

In our experiments we used \emph{Debian GNU/Linux 7.2}, \emph{Python 2.7} 
and an \emph{Intel Core i7-3770 3.4 GHz} processor. Timing measurements 
were made on a single dedicated core with garbage collection disabled. 
The Git repository for this 
paper\footnote{\url{https://github.com/OlivierBlanvillain/bfc-paper}} 
contains the necessary scripts and instructions to reproduce all the 
evaluation experiments presented in this section. The crawler source 
code is available under the MIT license from the project's 
websites\footnote{\url{https://github.com/BlogForever/crawler}}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extraction success rates}

To evaluate article and title extraction from weblog posts 
we compare our approach to three open source projects: 
Readability\footnote{\url{https://github.com/gfxmonk/python-readability}}, 
Boilerpipe~\cite{kohlschuetter2010} and Goose\footnote{\url{https://github.com/GravityLabs/goose}}, 
which are implemented in JavaScript, Java and Scala respectively. 
These projects are more generic than our blog-specific approach in the 
sense that they are able to identify and extract data directly from HTML 
source code, and do not make use of web feeds or structural similarities 
between pages of the same weblog (observations (\ref{havefeedAssum}) and 
(\ref{similarhtmlAssum})). Table \ref{table:evaluation-performance} shows the extraction 
success rates for article and title on a test sample of 2300 posts 
from 230 weblogs obtained from the Spinn3r dataset~\cite{burton2011}.

\begin{table}
\tbl{Extraction success rates for different algorithms.}
{\begin{tabular}{@{}lcccc@{}} \toprule
		\textbf{Target} & \textbf{Our~approach} & \textbf{Readability} & \textbf{Boilerpipe} & \textbf{Goose} \\ \hline
    Article & 93.0\% & 88.1\% & 79.3\% & 79.2\% \\
    Title   & 95.0\% & 74.0\% & N/A    & 84.9\% \\ \hline
\end{tabular}}
\label{table:evaluation-performance}
\end{table}

On our test dataset, \autoref{extractionAlgo} outperformed the 
competition by 4.9\% on article extraction and 10.1\% on title 
extraction. It is important to stress that Readability, Boilerpipe 
and Goose rely on generic techniques such as word density, paragraph 
clustering and heuristics on HTML tagging conventions, which are designed 
to work for any type of web page. On the contrary, our algorithm is 
only suitable for pages with associated web feeds, as these provide the 
reference data used to build extraction rules. Therefore, results shown 
in Table \ref{table:evaluation-performance} should not be interpreted as a general 
quality evaluation of the different projects, but simply as evidence 
that our approach is more suitable when working with weblogs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Article extraction running times}

In addition to the quality of the extracted data we also evaluated the 
running time of the extraction procedure. The main point of interest is 
the ability of the extraction procedure to scale as the number of posts 
in the processed weblog increases. This corresponds to the evaluation of 
a \emph{NewCrawl} task, which is in charge of harvesting all published 
content on a weblog.

Figure \ref{runningtime} shows the cumulated time spent for each article 
extraction procedure (this excludes common tasks such as downloading 
pages and storing results) as a function of the number of weblog posts 
processed. We used the Quantum Diaries\footnote{\url{http://www.quantumdiaries.org}} 
blog for this experiment. Data presented in this graph was obtained
by taking the arithmetic mean over 10 measurements. These results are
believed to be significant given that standard deviations are of
the order of 2 milliseconds.

\begin{figure*}[!hbt]
\centering
\label{runningtime}
\input{main-gnuplottex-fig1.tex}
\end{figure*}

As illustrated in \autoref{runningtime}, our approach spends the majority 
of its total running time between the initialisation the processing 
of the first weblog post. This initial increase of about 0.4 seconds 
corresponds to cost of executing \autoref{linearAlgo} to compute 
extraction rule for articles. As already mentioned, this consists 
in computing the \emph{best extraction rule} of each pages references 
by the web feed and picking the one functioning best on these pages. 
Once we have this extraction rule, processing subsequent weblog posts only 
requires parsing and applying the rule, which takes about 3 milliseconds 
and are barely visible on the scale of \autoref{runningtime}. The other 
evaluated solutions do not function this way: each weblog post is processed 
as new and independent input, leading to approximately linear running 
times.

The vertical dashed line at 15 processed weblog posts represents a suitable 
point of comparison of processing time per post. Indeed, as the 
web feed of our test weblog contains 15 posts, the extraction rule 
computation performed by our approach include the cost of entirely 
processing these 15 entries. That being said, comparing raw performance 
of algorithms implemented in different programming languages is not very 
informative given the high variations of running times observed across 
programming languages~\cite{hundt2011}.